Assessing the spatial patterning of Middle Paleolithic human settlement in Western Iberia

### 2020-10-30 analysis/paper/paper.Rmd
added alentejo sites

library(rgdal)
library(raster)
library(rasterVis)
library(lattice)
library(foreach)
library(rnaturalearth)
library(tidyverse)
library(ggplot2)
library(broom)
library(foreach)
library(sf)
library(osfr)
library(cowplot)
library(gridExtra)
library(rela)
library(psych)
library(factoextra)
library(FactoMineR)

@@

#Convert coordinates to longlat WGS84
sites_sub_long <- spTransform(sites_sub, "+proj=longlat +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")

caves_long <- spTransform(caves, "+proj=longlat +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")

@@

# Extract mean values from stack for each site and all caves within a 20 m radius
sites_vals <- raster::extract(terrainstack,
                      sites_sub_long,
                      buffer = 20,
                      fun = mean,
                      sp = TRUE)

@@

rasterVis::levelplot(areaDEM,
                     margin = list(x = FALSE,
                                   y = TRUE),
                     col.regions = terrain.colors(16),
                     xlab = list(label = "",
                                 vjust = -0.25),
                     sub = list(
                       label = "Meters a.s.l.",
                       font = 1,
                       cex = .9,
                       hjust = 0.5),
                     colorkey=list(space="bottom"),
                     key = list(
                       space = "left",
                       points = list(
                         pch = c(18,20),
                         col = c("red","blue")),
                       text = list(
                         c("Cave","Open-air"),
                         cex=1))) +
  spplot(sites_sub_long, # add a layer of points
         zcol = "Type",
         cex = 2,
         pch = c(18,20),
         col.regions = c("red","blue"))

@@

pca_data_active2 <- dplyr::select(pca_data, Elevation,
                                 Slope, Aspect, Distance, Latitude)

res.pca <- PCA(pca_data_active2, scale = TRUE, graph = F, quanti.sup = c(5))

groups <- as.factor(pca_data$Dmn_R_M)
groups_2 <- as.factor(pca_data$Type)

>>>>

library(rgdal)
library(raster)
library(rasterVis)
library(lattice)
library(foreach)
library(rnaturalearth)
library(tidyverse)
library(ggplot2)
library(broom)
library(foreach)
library(sf)
library(osfr)
library(cowplot)
library(gridExtra)
library(rela)
library(psych)
library(factoextra)
library(FactoMineR)
library(tab)

@@

@@

# Extract mean values from stack for each site and all caves within a 20 m radius
sites_vals <- raster::extract(terrainstack,
                      sites_sub,
                      buffer = 20,
                      fun = mean,
                      sp = TRUE)

@@

rasterVis::levelplot(areaDEM,
                     margin = list(x = FALSE,
                                   y = TRUE),
                     col.regions = terrain.colors(16),
                     xlab = list(label = "",
                                 vjust = -0.25),
                     sub = list(
                       label = "Meters a.s.l.",
                       font = 1,
                       cex = .9,
                       hjust = 0.5),
                     colorkey=list(space="bottom"),
                     key = list(
                       space = "left",
                       points = list(
                         pch = c(18,20),
                         col = c("red","blue")),
                       text = list(
                         c("Cave","Open-air"),
                         cex=1))) +
  spplot(sites_sub, # add a layer of points
         zcol = "Type",
         cex = 2,
         pch = c(18,20),
         col.regions = c("red","blue"))

@@

pca_data_active2 <- pca_data %>%  
  drop_na() %>% 
  dplyr::select(Elevation, Slope, Aspect, Distance, Latitude)

res.pca <- PCA(pca_data_active2, scale = TRUE, graph = F, quanti.sup = 5)

groups <- pca_data %>% 
  dplyr::select(Dmn_R_M) %>% 
  drop_na()
groups <- as.factor(groups$Dmn_R_M)

groups_2 <- as.factor(pca_data$Type)
